---
title: "Untitled"
author: "Jun Ming Li"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries used
```{r}

library(car)
library(leaps)
library(olsrr)
library(readr)
library(dplyr)
library(ggplot2)
library(readxl)

```



```{r}

# Load and preview Data
prev_app_Final         <- read.csv("previous_application_final_cleaned_17VARS.csv")
# prev_app_Num           <- read.csv("previous_application_numvars_cleaned.csv")
appdata_final          <- read_excel("application_data_cleaned_ALL_FILES.xlsx", sheet = 3)

# appdata_numvars        <- read_excel("application_data_cleaned.xlsx", sheet = 1)
 excel_sheets("application_data_cleaned_ALL_FILES.xlsx")



# str(prev_app_Final)
# str(appdata_final)
# 



```






# Join both data and match to have the same ID: appdata_final$SK_ID_CURR  = prev_app_Final$SK_ID_CURR 

```{r}



# inner join, so that we only have columns where the IDs match 
# https://www.youtube.com/watch?v=Yg-pNqzDuN4 


# Rename all previous columns with a suffix
prev_app_Final <- prev_app_Final %>% rename_with(~ paste(., "prev", sep = "_"))
prev_app_Final <- rename(prev_app_Final, X =X_prev, SK_ID_CURR = SK_ID_CURR_prev)


Merged_data <- inner_join(appdata_final,prev_app_Final, by='SK_ID_CURR', suffix = c("_df1", "_df2"))

# df1 = appdata_final
# df2 = prev_app_Final


## Check to make sure the suffix naming convention works

# # items from appdata_final
# Merged_data %>% select(contains('df1'))
# 
# # items from prev_app_Final
# Merged_data %>% select(contains('df2'))


str(Merged_data)
head(Merged_data)

```





# Mazen's ideas

```{r}


# ratios of loan amount given out to price of item being bought with the loan
Merged_data$LoanGoodR_df1 <- Merged_data$AMT_CREDIT/Merged_data$AMT_GOODS_PRICE
Merged_data$LoanGoodR_df2 <- Merged_data$AMT_CREDIT_prev/Merged_data$AMT_GOODS_PRICE_prev
# This is officially called the Loan to Value ratio



# check for any NA values (where loan given out is 0)
sum(is.na(Merged_data$LoanGoodR_df1))

sum((is.na(Merged_data$LoanGoodR_df2)))

which((is.na(Merged_data$LoanGoodR_df2)))
Merged_data$LoanGoodR_df2[11]


# find the frequency of ratios
LoanstoGooddf1_frequency <- data.frame(sort(table(Merged_data$LoanGoodR_df1), decreasing = TRUE))
colnames(LoanstoGooddf1_frequency) = c("Ratio","Count") #changing the column names for easier understanding
LoanstoGooddf1_frequency


# plot the distribution
plot(table(Merged_data$LoanGoodR_df1), ylab="Frequency Count", xlab = "Ratio of Loan to Good")


```




```{r}

plot(table(Merged_data$REGION_POPULATION_RELATIVE))

ggplot(Merged_data,aes(x=REGION_POPULATION_RELATIVE)) +geom_histogram()

# The bins seem to be from 0 to .1 , .1 to .2 , and .2 to .3, and .3+
# The data is mostly symmetrical, with only a few outlier points. 


# If we find outliers (from the residual plot) in our regression model, we can change the outliers into the mean/median/mode of the value in the columns.and we-add them back in




```




# Dimension Reduction of non-numeric data



```{r}

# Find variables that are factors
# Merged_data %>% select(!is.numeric)

# Education
unique(Merged_data$NAME_EDUCATION_TYPE) #Change to Middle, HS, College+. The "Incomplete higher" will be classified as college 

Merged_data$NAME_EDUCATION_TYPE <- case_when(Merged_data$NAME_EDUCATION_TYPE %in% c("Lower secondary") ~ "Middle School", .default = Merged_data$NAME_EDUCATION_TYPE)

Merged_data$NAME_EDUCATION_TYPE <- case_when(Merged_data$NAME_EDUCATION_TYPE %in% c("Secondary / secondary special") ~ "High School", .default = Merged_data$NAME_EDUCATION_TYPE)

Merged_data$NAME_EDUCATION_TYPE <- case_when(Merged_data$NAME_EDUCATION_TYPE %in% c("Higher education", "Incomplete higher","Academic degree" ) ~ "College+", .default = Merged_data$NAME_EDUCATION_TYPE)



# Employment status
unique(Merged_data$NAME_INCOME_TYPE)    #Unemployed, Retired, Student, Employed
Merged_data$NAME_INCOME_TYPE <- case_when(Merged_data$NAME_INCOME_TYPE %in% c("Working","State servant","Commercial associate", "Maternity leave")~ "Employed", .default = Merged_data$NAME_INCOME_TYPE)

Merged_data$NAME_INCOME_TYPE <- case_when(Merged_data$NAME_INCOME_TYPE %in% c("Pensioner") ~ "Retired", .default = Merged_data$NAME_INCOME_TYPE)

       
# Marital Status       
unique(Merged_data$NAME_FAMILY_STATUS) #Married, Widowed, or Single
Merged_data$NAME_FAMILY_STATUS <- case_when(Merged_data$NAME_FAMILY_STATUS %in% c("Separated", "Single") ~ "Single", .default = Merged_data$NAME_FAMILY_STATUS)
Merged_data$NAME_FAMILY_STATUS <- case_when(Merged_data$NAME_FAMILY_STATUS %in% c("Married", "Civil marriage") ~ "Married", .default = Merged_data$NAME_FAMILY_STATUS)


# Housing Type
unique(Merged_data$NAME_HOUSING_TYPE)

# municipal housing is government provided housing, group with rented apartment.
# coop apartment, they own a part of the building (usually a room if they can afford the shares for one whole room), group with housing/apartment
# office apartmenmt is a building made by a corporation and workers can rent it out to live close by

Merged_data$NAME_HOUSING_TYPE <- case_when(Merged_data$NAME_HOUSING_TYPE %in% c("House / apartment","Co-op apartment"  ) ~ "Owned", .default = Merged_data$NAME_HOUSING_TYPE)

Merged_data$NAME_HOUSING_TYPE <- case_when(Merged_data$NAME_HOUSING_TYPE %in% c("Rented apartment","Municipal apartment", "Office apartment") ~ "Renting", .default = Merged_data$NAME_HOUSING_TYPE)

Merged_data$NAME_HOUSING_TYPE <- case_when(Merged_data$NAME_HOUSING_TYPE %in% c("With parents" ) ~ "Doesn't Own", .default = Merged_data$NAME_HOUSING_TYPE)



# Finished by Jairo already
unique(Merged_data$OCCUPATION_TYPE)


# Organization type (We have decided to remove this variable from our study)
unique(Merged_data$ORGANIZATION_TYPE)
# https://jccc.libguides.com/c.php?g=511520&p=3494606 To define industry type

Merged_data$ORGANIZATION_TYPE <- NULL



# Contract type, already in 2 categories
unique(Merged_data$NAME_CONTRACT_TYPE)

# Jairo Finished
unique(Merged_data$NAME_CONTRACT_TYPE_prev)



unique(Merged_data$NAME_CONTRACT_STATUS_prev)
# Canceled, either the client or bank stopped the loan application process
# Refused, by the bank
# Unused, the loan was granted, but did not use the credit line

unique(Merged_data$NAME_CASH_LOAN_PURPOSE_prev)
unique(Merged_data$NAME_CONTRACT_STATUS_prev)
unique(Merged_data$NAME_GOODS_CATEGORY_prev)



```

# Dimension Reduction of numeric data

```{r}

# Jairo fixed the rest in his data clean up code

```



## Make the categorical variables into factors

```{r}
str(Merged_data)


# Non-numerical ones

categorical_names <- names(dplyr::select(Merged_data,
  TARGET,
  CNT_CHILDREN, #do not remove this. This makes the AUC below .5
  CODE_GENDER,
  FLAG_OWN_CAR,
  FLAG_OWN_REALTY,
  NAME_INCOME_TYPE,
  NAME_EDUCATION_TYPE,
  NAME_FAMILY_STATUS,
  NAME_HOUSING_TYPE,
  OCCUPATION_TYPE,
  WEEKDAY_APPR_PROCESS_START,
  NAME_CONTRACT_STATUS_prev,
  NAME_CONTRACT_TYPE,
  NAME_CASH_LOAN_PURPOSE_prev,
  # NAME_PAYMENT_TYPE_prev, # has NAs, must remove it
  NAME_GOODS_CATEGORY_prev,
  NAME_YIELD_GROUP_prev,
  NFLAG_INSURED_ON_APPROVAL_prev,
  NAME_CONTRACT_TYPE_prev,
  REG_CITY_NOT_LIVE_CITY
))

Merged_data[,categorical_names] <- lapply(Merged_data[,categorical_names], FUN= as.factor)
str(Merged_data)



unique(Merged_data$FLAG_EMP_PHONE)
unique(Merged_data$FLAG_CONT_MOBILE)
unique(Merged_data$FLAG_EMAIL)


# Numerical ones

categorical_names2 <- names(Merged_data %>% dplyr::select(
  FLAG_EMP_PHONE,
  FLAG_CONT_MOBILE,
  FLAG_EMAIL,
  FLAG_DOCUMENT_2
))

Merged_data[,categorical_names2] <- lapply(Merged_data[,categorical_names2], FUN= as.factor)
str(Merged_data)

# Old Code
# categorical_names2 <- names(Merged_data %>% select(contains('FLAG')))
# Merged_data[,categorical_names2] <- lapply(Merged_data[,categorical_names2], FUN= as.factor)
# str(Merged_data)



#Final Check
str(Merged_data)






```



## Remove Unnecessary Variables

```{r}
str(Merged_data)

# remove identifier keys
Merged_data$X <- NULL
Merged_data$SK_ID_CURR <- NULL
Merged_data$SK_ID_PREV_prev <- NULL


# This data only has 1 entry value for all rows
unique(Merged_data$FLAG_MOBIL)
Merged_data$FLAG_MOBIL <- NULL

Merged_data$NAME_PAYMENT_TYPE_prev <- NULL
Merged_data$DEF_60_CNT_SOCIAL_CIRCLE <- NULL

```



## check for values that are 0 and change them to infinitly small ones. Also check for Inf values for our ratios.

```{r include=FALSE}
str(Merged_data)


# LoanValue ratio
Merged_data$LoanGoodR_df1[Merged_data$LoanGoodR_df1 == 0]
Merged_data$LoanGoodR_df1[which(is.na(Merged_data$LoanGoodR_df1))]
Merged_data$LoanGoodR_df1[which(is.infinite(Merged_data$LoanGoodR_df1))] <- .999

# LoanValue ratio previous application
Merged_data$LoanGoodR_df2[Merged_data$LoanGoodR_df2 == 0]
Merged_data$LoanGoodR_df2[which(is.na(Merged_data$LoanGoodR_df2))] <- .001
Merged_data$LoanGoodR_df2[which(is.infinite(Merged_data$LoanGoodR_df2))] <- .999


# AMT_INCOME_TOTAL
Merged_data$AMT_INCOME_TOTAL[Merged_data$AMT_INCOME_TOTAL == 0]
Merged_data$AMT_INCOME_TOTAL[which(is.na(Merged_data$AMT_INCOME_TOTAL))] <- .001


# AMT_ANNUITY 
Merged_data$AMT_ANNUITY[Merged_data$AMT_ANNUITY == 0]
Merged_data$AMT_ANNUITY[which(is.na(Merged_data$AMT_ANNUITY))] <- .001

# AMT_GOODS_PRICE
Merged_data$AMT_GOODS_PRICE[Merged_data$AMT_GOODS_PRICE == 0]
Merged_data$AMT_GOODS_PRICE[which(is.na(Merged_data$AMT_GOODS_PRICE))] <- .001

# REGION_POPULATION_RELATIVE
Merged_data$REGION_POPULATION_RELATIVE[Merged_data$REGION_POPULATION_RELATIVE == 0]
Merged_data$REGION_POPULATION_RELATIVE[which(is.na(Merged_data$REGION_POPULATION_RELATIVE))] <- .001


# DAYS_EMPLOYED
Merged_data$DAYS_EMPLOYED[Merged_data$DAYS_EMPLOYED == 0] <- .001
Merged_data$DAYS_EMPLOYED[which(is.na(Merged_data$DAYS_EMPLOYED))]

# DAYS_REGISTRATION
Merged_data$DAYS_REGISTRATION[Merged_data$DAYS_REGISTRATION == 0] <- .001
Merged_data$DAYS_REGISTRATION[which(is.na(Merged_data$DAYS_REGISTRATION))]

# DAYS_ID_PUBLISH 
Merged_data$DAYS_ID_PUBLISH[Merged_data$DAYS_ID_PUBLISH == 0] <- .001
Merged_data$DAYS_ID_PUBLISH[which(is.na(Merged_data$DAYS_ID_PUBLISH))]

# APARTMENTS_AVG
Merged_data$APARTMENTS_AVG[Merged_data$APARTMENTS_AVG == 0] <- .001
Merged_data$APARTMENTS_AVG[which(is.na(Merged_data$APARTMENTS_AVG))]


# LANDAREA_AVG 
Merged_data$APARTMENTS_AVG[Merged_data$APARTMENTS_AVG == 0] <- .001
Merged_data$APARTMENTS_AVG[which(is.na(Merged_data$APARTMENTS_AVG))]

# DAYS_LAST_PHONE_CHANGE  
Merged_data$DAYS_LAST_PHONE_CHANGE[Merged_data$DAYS_LAST_PHONE_CHANGE == 0] <- .001
Merged_data$DAYS_LAST_PHONE_CHANGE[which(is.na(Merged_data$DAYS_LAST_PHONE_CHANGE))]


# AMT_REQ_CREDIT_BUREAU_HOUR
Merged_data$AMT_REQ_CREDIT_BUREAU_HOUR[Merged_data$AMT_REQ_CREDIT_BUREAU_HOUR == 0] <- .001
Merged_data$AMT_REQ_CREDIT_BUREAU_HOUR[which(is.na(Merged_data$AMT_REQ_CREDIT_BUREAU_HOUR))]

# AMT_ANNUITY_prev
Merged_data$AMT_ANNUITY_prev[Merged_data$AMT_ANNUITY_prev == 0] <- .001
Merged_data$AMT_ANNUITY_prev[which(is.na(Merged_data$AMT_ANNUITY_prev))] <- .001

# AMT_APPLICATION_prev
Merged_data$AMT_APPLICATION_prev[Merged_data$AMT_APPLICATION_prev == 0] <- .001
Merged_data$AMT_APPLICATION_prev[which(is.na(Merged_data$AMT_APPLICATION_prev))] 

# AMT_CREDIT_prev
Merged_data$AMT_CREDIT_prev[Merged_data$AMT_CREDIT_prev == 0] <- .001
Merged_data$AMT_CREDIT_prev[which(is.na(Merged_data$AMT_CREDIT_prev))] <- .001

# DAYS_DECISION_prev 
Merged_data$DAYS_DECISION_prev[Merged_data$DAYS_DECISION_prev == 0] <- .001
Merged_data$DAYS_DECISION_prev[which(is.na(Merged_data$DAYS_DECISION_prev))] <- .001

# CNT_PAYMENT_prev
Merged_data$CNT_PAYMENT_prev[Merged_data$CNT_PAYMENT_prev == 0] <- .001
Merged_data$CNT_PAYMENT_prev[which(is.na(Merged_data$CNT_PAYMENT_prev))] <- .001


# DAYS_FIRST_DRAWING_prev
Merged_data$DAYS_FIRST_DRAWING_prev[Merged_data$DAYS_FIRST_DRAWING_prev == 0] <- .001
Merged_data$DAYS_FIRST_DRAWING_prev[which(is.na(Merged_data$DAYS_FIRST_DRAWING_prev))] <- .001


# Replace NAs with the mean of Amt goods Price from previous application

Merged_data$AMT_GOODS_PRICE_prev[which(is.na(Merged_data$AMT_GOODS_PRICE_prev))] <- mean(Merged_data$AMT_GOODS_PRICE_prev, na.rm = TRUE)

# sapply(Merged_data, function(x) which(is.na(x)))

```





## Further manual review and removal of variables 
```{r}

# str(Merged_data)
# 
# Merged_data %>% select(is.numeric)

# children count, has 7 dummy variables...either convert to numeric or remove...I want to remove it for now
# weekday, has 7 dummy variables, the day was not that important when we discussed it)
# REG_CITY_NOT_LIVE_CITY, their vicinity or how close to a city may not seem important for now
# NAME_GOODS_CATEGORY_prev, because NAME_CASH_LOAN_PURPOSE_prev sort of already covers "Why" or "what" the loan is used for

# # 2nd run removal

# 
# 
Merged_data$CNT_CHILDREN <- NULL
Merged_data$WEEKDAY_APPR_PROCESS_START <- NULL
Merged_data$REG_CITY_NOT_LIVE_CITY <- NULL
Merged_data$NAME_GOODS_CATEGORY_prev <- NULL
Merged_data$NAME_INCOME_TYPE <- NULL


Merged_data$FLAG_EMP_PHONE <- NULL
Merged_data$FLAG_CONT_MOBILE <- NULL
Merged_data$DAYS_LAST_PHONE_CHANGE <- NULL

Merged_data$NAME_FAMILY_STATUS <- NULL
Merged_data$FLAG_EMAIL <- NULL
Merged_data$APARTMENTS_AVG <- NULL
Merged_data$LANDAREA_AVG <- NULL

# Merged_data$REGION_POPULATION_RELATIVE
Merged_data$AMT_ANNUITY <- NULL
Merged_data$AMT_ANNUITY_prev <- NULL
Merged_data$AMT_REQ_CREDIT_BUREAU_HOUR <- NULL
Merged_data$DAYS_FIRST_DRAWING_prev <- NULL
Merged_data$NAME_YIELD_GROUP_prev <- NULL
Merged_data$NFLAG_INSURED_ON_APPROVAL_prev <- NULL
Merged_data$AMT_APPLICATION_prev <- NULL
Merged_data$DAYS_ID_PUBLISH <- NULL
Merged_data$DAYS_REGISTRATION <- NULL

# 
# # remvoval after forward stepwise
# 
Merged_data$LoanGoodR_df2 <- NULL
Merged_data$NAME_CASH_LOAN_PURPOSE_prev <-  NULL
Merged_data$CODE_GENDER <- NULL
Merged_data$AMT_GOODS_PRICE <- NULL
Merged_data$AMT_GOODS_PRICE_prev <- NULL
Merged_data$AMT_CREDIT <- NULL
Merged_data$AMT_CREDIT_prev <- NULL
Merged_data$CNT_PAYMENT_prev <- NULL

str(Merged_data)

# NAME_CONTRACT_STATUS showed that only one of the levels seem to be significant, but due to how important of a metric this generally is , we have decided to keep this in our model. 


# LoanGoodR_df2 was considered not significant, which shows that the current application's metrics are weighted a bit more. 

# NAME_CASH_LOAN_PURPOSE_prev, not statistically significant
# Gender- bank cannot discriminate on gender.

 
# The above variables were further removed from our initial data reduction after more literature review and consideration of the typical metric by which banks judge a person's credit worthiness. Moreover, variables suchs as gender, were removed because banks cannot feasible enforce restrictions against those variables under penalty of law. 

```


```{r}

summary(Merged_data)


```

## Summary Statistics generation

```{r}

str(Merged_data)


multiple.func <- function(x) {
      c(min = min(x), mean = mean(x), max = max(x))
}

multiple.func2 <- function(x) {
      c(counts= table(x))
}


summary_num_stat <- as.data.frame(sapply(Merged_data %>% dplyr::select(where(is.numeric)), multiple.func))

summary_num_stat$Statistic <- c('Min','Mean','Max')
summary_num_stat <- summary_num_stat %>% relocate(Statistic, .after = last_col())

summary_factor_stats <- c(

Target = table(Merged_data$TARGET),
NAME_CONTRACT_TYPE = table(Merged_data$NAME_CONTRACT_TYPE),
FLAG_OWN_REALTY = table(Merged_data$FLAG_OWN_REALTY),
FLAG_OWN_CAR = table(Merged_data$FLAG_OWN_CAR),
NAME_EDUCATION_TYPE = table(Merged_data$NAME_EDUCATION_TYPE),
NAME_HOUSING_TYPE = table(Merged_data$NAME_HOUSING_TYPE),
OCCUPATION_TYPE = table(Merged_data$OCCUPATION_TYPE),
FLAG_DOCUMENT_2 = table(Merged_data$FLAG_DOCUMENT_2),
NAME_CONTRACT_TYPE_prev = table(Merged_data$NAME_CONTRACT_TYPE_prev),
NAME_CONTRACT_STATUS_prev = table(Merged_data$NAME_CONTRACT_STATUS_prev)

)

summary_factor_stats <- as.data.frame(summary_factor_stats)

library(writexl)


write_xlsx(x = list('Summary num stats' = summary_num_stat,
                    'summary factor stats' = summary_factor_stats
                    ), 'LoanDefaulterSummaryStatistics.xlsx', col_names = TRUE, format_headers = TRUE)


mean(Merged_data$AMT_INCOME_TOTAL)
mean(Merged_data$DAYS_BIRTH)
mean(Merged_data$DAYS_EMPLOYED)
mean(Merged_data$NAME_HOUSING_TYPE)
mean(Merged_data$FLAG_OWN_CAR)
mean(Merged_data$FLAG_OWN_REALTY)

```








## Partition data

```{r}

set.seed(123)

# Partition data into training (70%) and testing (30%) data sets

# create a sample of indexes
Merged_data.rows <- nrow(Merged_data)
Merged_data.index <- sample(Merged_data.rows, .7*Merged_data.rows)

# create datasets using above randomly chosen indexes
Merged_data.train <- Merged_data[Merged_data.index,]
Merged_data.test  <- Merged_data[-Merged_data.index,]

# confirm the total number of rows matches the above
nrow(Merged_data.test) + nrow(Merged_data.train)


```



## Balance the data

```{r}

# str(Merged_data.train)

table(Merged_data.train$TARGET)
table(Merged_data.train$TARGET)/nrow(Merged_data.train)

#      0      1 
#   903996  85594 
#  91.3%      8.7%


# 0 = no payment difficulties (this is the reference leve1)
# 1 = client with payment difficulties: he/she had late payment more than X days on at least one of the first Y installments of the loan in our sample



library(ROSE)

# balanced data set with over-sampling
# We are taking the minority class, and oversampling it. 
Merged_data.train.balanced.over <- ovun.sample(TARGET ~ ., data=Merged_data.train, p=0.5, method= "under")
Merged_data.train.balanced.over <- Merged_data.train.balanced.over$data


# Save the balanced dataset

Merged_data.train <- Merged_data.train.balanced.over
(table(Merged_data.train.balanced.over$TARGET)) / nrow(Merged_data.train)
```





## Scale the Train Dataset

```{r}
### Using Z-score

str(Merged_data.train)

# Standardize input variables

# create function to divide the days into year
divide365 <- function(x, na.rm = FALSE) round((x/365),2)

# creating my own scaling function because the R one produces errors when trying to predict or create a LM 
myscale <- function(x) (x - mean(x)) / sd(x)


Merged_data.train <- Merged_data.train %>% mutate(across(colnames(Merged_data.train %>% dplyr::select(contains('DAYS'))),divide365))

Merged_data.train <- Merged_data.train %>% mutate(across(colnames(Merged_data.train %>% dplyr::select(is.numeric)), myscale))


# str(Merged_data)


```




## Scale the Test Dataset

```{r}
### Using Z-score

str(Merged_data.test)

# Standardize input variables

# create function to divide the days into year
divide365 <- function(x, na.rm = FALSE) round((x/365),2)

# creating my own scaling function because the R one produces errors when trying to predict or create a LM 
myscale <- function(x) (x - mean(x)) / sd(x)


Merged_data.test <- Merged_data.test %>% mutate(across(colnames(Merged_data.test %>% dplyr::select(contains('DAYS'))),divide365))

Merged_data.test <- Merged_data.test %>% mutate(across(colnames(Merged_data.test %>% dplyr::select(is.numeric)), myscale))

# str(Merged_data)


```





## Logistic Regression, Full Model

```{r}



logit.reg.all <- glm(TARGET ~., data =  Merged_data.train, family = 'binomial')
summary(logit.reg.all)

str(Merged_data.train)

names(Merged_data.train)

```



```{r}


summary_model <- data.frame(summary(logit.reg.all)$coef)

summary_model$Index <- row.names(summary(logit.reg.all)$coef)

library(writexl)


write_xlsx(x = list( 
                     'summary_model' = summary_model
                     ), 'Logistic_Coef.xlsx', col_names = TRUE, format_headers = TRUE)




```






# Testing the Prediction on Testing Data Set
[Review this to make sure logistic also works](https://www.sthda.com/english/articles/36-classification-methods-essentials/150-stepwise-logistic-regression-essentials-in-r/)

```{r}


# General probabilities from logit model
predicted.test.prob <- predict(logit.reg.all, newdata=Merged_data.test, type = "response")


# We need to consider what the reference level of our target variable.
Merged_data.test$TARGET <- as.factor(Merged_data.test$TARGET)
Merged_data.train$TARGET <- as.factor(Merged_data.train$TARGET)
# The reference level is 0, meaning they didn't default. 

levels(Merged_data.test$TARGET)

# Use cutoff point of 0.5 to convert probabilities to class
predicted.test.class <- ifelse(predicted.test.prob > 0.5, "1", "0")



# Generating a confusion matrix
library(caret)
confusionMatrix(data = as.factor(predicted.test.class), reference = Merged_data.test$TARGET)


# sensitivity is the accuracy of predicting the positive class. 
# specificity is the accuracy of predicting the negative class. 



# Checking the AUC
library(pROC)
roc = roc(Merged_data.test$TARGET, predicted.test.prob)
auc(roc)
plot(roc)



```




## Variable selection 

### Forward selection 

```{r}

# Define base intercept only model
model.null <- glm(TARGET ~1, data =  Merged_data.train, family = 'binomial')


# Full model with all predictors
model.full <- logit.reg.all

library(MASS)


# Perform forward algorithm
reduced.forward.model <- stepAIC(model.null, scope =list(lower = model.null, upper = model.full), direction = "forward", trace = FALSE)
summary(reduced.forward.model)

forwardmodelsaved <- reduced.forward.model
summary(forwardmodelsaved)


```


